---
title: "time series patterns/graphics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(feasts)
library(tsibble)
library(readr)
library(tidyverse)
library(lubridate)
library(tsbox)
library(dygraphs)

admission_hourly <- read_rds("data/admission_hourly.rds")
admission_daily <- read_rds("data/admission_daily.rds")
```

# Lab 2: time series patterns/graphics

## time plot

You can create time plot using `autoplot()` function

```{r ts-plot}
??? |> ???(???) # create a time plot of daily data
??? |> ???() # create a time plot of hourly data
```

You can use filter_index or head() and tail() to select a subset of time series , try ?filter_index or google it


```{r filter-index}
admission_daily |> tsibble::???(???) |> autoplot(arrival) # plot only February 2016 
admission_daily |> tsibble::???(???) |> autoplot(arrival) # plot from February 2016 till the end
admission_daily |> tsibble::???(???) |> autoplot(arrival) # plot from February 2016 till the end
admission_hourly |> tsibble::filter_index(???) |> autoplot(arrival) # plot only 2015-01-01
admission_daily |> ???(???) |> autoplot() # first 100 observations
admission_daily |> ???(???) |> autoplot()  # last 100 observations
```

It could be really hard to visualise hourly or daily data, one possibility is to view a portion of the data, you can also use  interactive plots:

```{r admission-dygraph}
tsbox::ts_xts(admission_daily) |> 
  dygraph() |> 
  dyRangeSelector(dateWindow = c("2010-01-01", "2010-02-01"))

tsbox::ts_xts(admission_hourly) |> 
  dygraph() |> 
  dyRangeSelector(dateWindow = c("2010-01-01", "2010-01-02"))
```

You can also plot monthly, weekly, quarterly time series using ggplot2 package, see below for an example

```{r monthly-plot}
  monthly_admissions |> 
  autoplot(arrival) +
  labs(y = "arrival", x="Month",
       title = "Monthly A&E arrival",
       subtitle = "UK hospital")
```

## Seasonal plots

use seasonal and subseries plots to check whether series contain seasonality 

```{r seasonal-plot}
admission_daily |> feasts::???(???) #seasonal plot
admission_daily |> feasts::???(???)# subseries plot
```

You can also try this with hourly series, you can change the "period= ..." argument to see different types of seasonality

```{r multiple-seasonality}
??? |> feasts::???(arrival, ???) 
??? |> feasts::???(arrival, ???)

??? |> feasts::???(arrival, ???) 
??? |> feasts::???(arrival, ???)
```

Is there any seasonality in the daily time series? what about hourly and monthly?


How do you create a seasonal plot for the weekly, monthly and quarterly series series

```{r weekly-seasonality}
weekly_admissions <- admission_daily |> 
  index_by(week = yearweek(date)) |>
  summarise(arrival = sum(arrival))
???(weekly_admissions)
???(weekly_admissions)
```

```{r monthly-seasonality}
monthly_admissions <- admission_daily |> 
  index_by(month = yearmonth(date)) |>
  summarise(arrival = sum(arrival))
???(monthly_admissions)
???(monthly_admissions)

```


```{r quarterly-seasonality}
quarterly_admissions <- admission_daily |> 
  index_by(quarter = yearquarter(date)) |>
  summarise(arrival = sum(arrival))
???(quarterly_admissions)
???(quarterly_admissions)
```

## autocorrelation plot

```{r acf-calcul}
admission_daily |>
  feasts::???(arrival, lags = ???, geom = "point")# create lag plots for 14 lags, from 1 to 14
admission_daily |> feasts::???(lag_max = ???)# compute autocorrelation function for 14 lags
```

plot the autocorrelation 

```{r acf-plot}
admission_hourly |> ???(lag_max = ???) |> ???()# plot acf
admission_daily |> ???(lag_max = ???) |> ???()# plot acf
```

You can have a time plot, acf and season plot in one single plot if you use `gg_tsdisplay()` function

```{r multiple-plot-, options}
admission_hourly |> ???()#  
admission_daily |> ???()#  
```

You can use ljung box to test whether ACF is significant, if p-value is small, << 0.05 then there is a significant autocorrelation:

```{r ljung-box-test}
admission_daily |> ???(???, ljung_box)
```

What autocorrelation will tell us? Which key features could be highlighted by ACF?

## Make any graph using ggplot2 (optional)

You can create any graph that helps you to better understand data! I recommend you to look at the distributions of your variables, try geom_boxplot() , geom_histogram() and geom_density() which are helpful to better understand the variations

Here I tried to see if arrival of males or females is different over the weekend comparing to the weekday

```{r weekend}
weekend_an_weekday <- admission_hourly |> group_by(gender) |> 
  summarise(arrival=sum(arrival)) |> 
  mutate(
    Date=lubridate::as_date(time),
    hour=lubridate::hour(time),
    Day = lubridate::wday(time, label = TRUE),
    Weekend = (Day %in% c("Sun", "Sat"))) |> 
  filter(gender =="female") 
weekend_an_weekday |> ggplot(aes(x = hour, y = arrival)) +
  geom_line(aes(group=Date)) +
  facet_grid(Weekend ~., scales="free_y")
```

  
  