---
title: "Untitled"
output: html_document
date: "2023-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp3)
library(tidyverse)
arrival_daily_keys <- read_rds("data/arrival_hourly_keys.rds")
```

# Lab 3: time series features

For the daily admissions time series with keys (with gender, injury), extract the strength of trend and seasonality

  *Do you see any useful insight?*

```{r label, options}
arrival_daily_keys |>
  features(arrival, feat_stl)
```

Use ``GGally::ggpairs()`` to look at the relationships between the STL-based features. You might wish to change `seasonal_peak_year` and `seasonal_trough_year` to factors.

```{r label, options}
 library(GGally)
tourism |>
  ???(???, ???) |>
  select(-Region, -State, -Purpose) |>
  mutate(
    seasonal_peak_year = factor(???),
    seasonal_trough_year = factor(???),
  ) |>
  ggpairs()
```
 
 
Which is the peak quarter for holidays in each state?

```{r label, options}
tourism |>
  ???(State) |>
  ???(Trips = ???(Trips)) |>
  ???(???, ???) |>
  select(State, seasonal_peak_year)
```
 

Use a feature-based approach to look for outlying series in `PBS`. What is unusual about the series you identify as outliers?

```{r label, options}
library(broom)

## Compute features
PBS_feat <- PBS |>
  ????(Cost, feature_set(pkgs = "feasts")) |>
  na.omit()

## Compute principal components
PBS_prcomp <- PBS_feat |>
  select(-Concession, -Type, -ATC1, -ATC2) |>
  ???(scale = TRUE) |>
  ???(PBS_feat)

## Plot the first two components
PBS_prcomp |>
  ggplot(aes(x = ???, y = ???)) +
  geom_point()

## Pull out most unusual series from first principal component
outliers <- PBS_prcomp |>
  filter(.fittedPC1 > 7)
outliers

## Visualise the unusual series
PBS |>
  semi_join(outliers, by = c("Concession", "Type", "ATC1", "ATC2")) |>
  autoplot(Cost) +
  facet_grid(vars(Concession, Type, ATC1, ATC2)) +
  labs(title = "Outlying time series in PC space")
```
